---
title: "predmachlearn-011 Project"
author: "Grzegorz Tkaczyk"
date: "Saturday, February 21, 2015"
output: html_document
---

This is Practical Machine Learning project, based on data for personal activity comming from  
http://groupware.les.inf.puc-rio.br/har 

## Read data from the files 

```{r cache=TRUE}
setwd("E://Projects//R//predmachlearn-011")
d <- read.csv("pml-training.csv")
k <- read.csv("pml-testing.csv")
```

## Filter data

Data has been filtered for records with ```new_window = "no"```.
Variables has been limited to those that are not ```NA``` in the ```pml-testing.csv``` file.

```{r}


library(randomForest)
library(caret)
library(dplyr)
library(tidyr)

dz <- filter(d, new_window == "no" )
dz <- select(dz
            ,classe
,user_name,roll_belt,pitch_belt,yaw_belt,total_accel_belt,gyros_belt_x,gyros_belt_y
,gyros_belt_z,accel_belt_x,accel_belt_y,accel_belt_z,magnet_belt_x,magnet_belt_y
,magnet_belt_z,roll_arm,pitch_arm,yaw_arm,total_accel_arm,gyros_arm_x,gyros_arm_y
,gyros_arm_z,accel_arm_x,accel_arm_y,accel_arm_z,magnet_arm_x,magnet_arm_y,magnet_arm_z
,roll_dumbbell,pitch_dumbbell,yaw_dumbbell,total_accel_dumbbell,gyros_dumbbell_x
,gyros_dumbbell_y,gyros_dumbbell_z,accel_dumbbell_x,accel_dumbbell_y,accel_dumbbell_z
,magnet_dumbbell_x,magnet_dumbbell_y,magnet_dumbbell_z,roll_forearm,pitch_forearm
,yaw_forearm,gyros_forearm_x,gyros_forearm_y,gyros_forearm_z,accel_forearm_x
,accel_forearm_y,accel_forearm_z,magnet_forearm_x,magnet_forearm_y,magnet_forearm_z
)
```

Split data set to da(60%) and db(40%) subsets. da will be used for training, da for testing.

```{r}
trainIndex = createDataPartition(dz$classe, p = 0.60,list=FALSE)
da = dz[trainIndex,]
db = dz[-trainIndex,]
```

## Use random forest model

Random forest model to build predition algorithm.

```{r}
modFit <- randomForest(formula = classe ~ ., data = da)
modFit
```
**OOB estimate of  error rate is 0.69%**

## Cross validate the model using db testing subset

```{r}
pred <- predict(modFit,db) 
db$predRight <- pred==db$classe
table(db$predRight)
table(pred,db$classe)
paste((1-sum(db$predRight)/length(db$predRight))/100 , "%")
```

**Estimated error after cross validation with testing subset is 0.59%.**

## Extract results to the files

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pred_k <- predict(modFit,k)
pml_write_files(pred_k)

```

Files has been submited. Score 20/20